\newglossaryentry{modelmatix}{name={model matrix}, description={An objects position, rotation and scale represented as matrix}, plural={model matrices}}

\newglossaryentry{cameramatrix}{name={camera matrix}, description={The cameras position and rotation represented as matrix. Other term for the camera's \gls{modelmatix}}.}

\newglossaryentry{viewmatrix}{name={view matrix}, description={A view matrix is used to draw from the viewpoint of the oject the viewmatrix belongs to.}, plural={view matrices}}



\newglossaryentry{portalpair}{name={portal pair}, description={Two connected portals. When travelling through a portal pair's portal the  object is teleported to the  portal pair's other portal.}}

\newglossaryentry{endpoint}{name={endpoint}, description={One portal of a portal pair.}}

\newglossaryentry{teleportationmatrix}{name={teleportation matrix}, description={An endpoint's matrix that when applied to an object moves that object to the other endpoints location}, plural={teleportation matrices}}

\newglossaryentry{renderpass}{name={renderpass}, description={A renderpass defined by Vulkan specification \cite{khronos:vulkan:spec1.1}. Can have multiple subpasses.}}




\newglossaryentry{subpass}{name={subpass}, description={A subpass defined by Vulkan specification \cite{khronos:vulkan:spec1.1}. Belongs to a renderpass.}}

\newglossaryentry{portalset}{name={portal set}, description={The set of all Portals.}}
\newglossaryentry{portalcount}{name={portal count}, description={The number of portals within the \gls{portalset}}}
\newglossaryentry{portalid}{name={portal set}, description={An unique integer indentifying a portal.}}

\newglossaryentry{object}{name={object}, description={A entity which can be drawn.}}
\newglossaryentry{objectset}{name={object set}, description={The set of all Objects}}

\newglossaryentry{recursioncount}{name={recursion count}, description={How often the view can traverse through portals.}}


\newglossaryentry{viewray}{name={view ray}, description={A ray that is cast for raytring. Its intersection determines a pixel's color.}}




